The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-24 paper.tex
tried to do the merging ... not sure everything went fine :)

Figure~\ref{roomba-perception} gives a first perspective on how naive
users perceive robots over a long period of time. Here, \emph{Roomba} vacuum cleaning robots
were given to nine households for a 6-months period. The aim was to explore how
people use and perceive the robot over time, and how they generally live with
the Roomba in their home. At various time points, users were asked to rate
their perception of different features of the robot on a 7-point rating scale. As
Figure~\ref{roomba-perception} shows, some features remained stable over time
(\emph{ease of use} and \emph{fun}), and others decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one does not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on his/her life. Even if one can argue that the Roomba is not designed
to foster interaction, the question is raised: how to sustain engagement between
a robot and a human over a long period of time? And, as a pre-requisite, do we
actually understand the psychological phenomenon, and can we effectively account
for the dynamics of these interactions?

@@

\fxnote{state here our main research hypothesis}
\fxnote{state here our main contributions}

@@

From the Roomba study, we can report on several findings related to
anthropomorphism and the anthropomorphic design of robots.
%The gathered data for these aspects include qualitative observations,
%semi-structured interviews, and also quantitative assessments of various
%aspects of human-robot interaction.
The Roomba study is one of very few long-term human-robot interaction studies in
the domestic environment.

In the following, we present findings in respect of two aspects: First, we
qualitatively describe what kinds of anthropomorphic interactions between
participants and the robot were observed and outline how these interactions
evolved over time.  Second, we present people's perception of human-likeness and
social capabilities of the robot and how this perception changed over
time.\footnote{\hl{note that these two different kinds of data
(behavior/interaction and perception) already show that we try to measure
anthropomorphism from these two angles!}}

@@

    \begin{tikzpicture}[scale=0.5, y = 2cm]
    %axis
    \draw (0,0) -- coordinate (x axis mid) (20,0);
    \draw (0,0) -- coordinate (y axis mid) (0,7);
    %ticks
    \foreach \x in {0,4,...,20}
        \draw (\x,1pt) -- (\x,-3pt);
    \foreach \y in {1,...,7} 
        \draw[black!20] (20,\y) -- (-3pt,\y);
    \foreach \y in {1,...,7} 
        \draw (-3pt, \y) node[anchor=east] {\y}; 
    %labels
    \node[below=0.8cm] at (x axis mid) {People's perception ($\sigma=15$)};
    \node at (2, -0.4) {\tiny expectations};
    \node[align=center] at (6, -0.4) {\tiny after\\ \tiny first usage};
    \node[align=center] at (10, -0.4) {\tiny after\\ \tiny 2 weeks};
    \node[align=center] at (14, -0.4) {\tiny after\\ \tiny 2 months};
    \node[align=center] at (18, -0.4) {\tiny after\\ \tiny 6 months};

    \draw[dashed, black!50] (6,0) -- (6,7) node[sloped, below, pos=0.25] {\tiny novelty effect};

    \draw[ultra thick, red!50] (2,5.6) -- (6,6.1) -- (10,6.3) -- (14,6.3) -- (18,6.1) node at (15,6.5) {\tiny \it ease of use};
    \draw[ultra thick, purple] (2,4.9) -- (6,5.2) -- (10,5.1) -- (14,4.5) -- (18,4.2) node at (3,4.7) {\tiny \it usefulness};
    \draw[ultra thick, blue!50] (2,4) -- (6,4.1) -- (10,4.6) -- (14,4.1) -- (18,4.3) node at (16,4) {\tiny \it fun};
    \draw[ultra thick, green!50] (2,3.9) -- (6,4.1) -- (10,4.2) -- (14,3.8) -- (18,3.3) node at (19,3.5) {\tiny \it intelligence};
    \draw[ultra thick, orange] (2,3.5) -- (6,4.2) -- (10,4) -- (14,3.2) -- (18,2.8) node at (14,2.7) {\tiny \it impact};

    \draw[ultra thick, dashed] (2,5.2) -- (6,5.2) -- (10,5) -- (14,4.9) -- (18,4.5) node at (14,5.3) {\tiny \it overall};

@@

In accordance with previous work we also found relatively low rates of users
treating and perceiving their vacuum cleaning robot as a living (or human-like)
entity \citep{sung_housewives_2008}. We interpret that it might be the pure
functional purpose and non human-like form of the Roomba that prevented people
from anthropomorphizing it. Still, we observed some initial anthropomorphic
effects, both reflected in interaction with and perception of the robot.
However, most of the anthropomorphic effects did not last long. \hl{One fairly
stable human-like interaction with the robot was \textit{talking to the robot},
despite the fact that the robot was not able to recognize and respond to verbal
statements.\footnote{important! I think I've seen that finding somewhere else
when comparing ASIMO to AIBO or so.}}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{aibo-roomba-ipad_2.jpg}
    \caption{The Sony AIBO robotic dog (left), the iRobot Roomba vacuum-cleaning robot (middle), and the Apple Inc. iPad tablet computer (right) were compared in terms of how far users of these devices use anthropomorphic expressions in discussions about their technology / robot in online discussion forums.}
    \label{fig:aibo-roomba-ipad}
\end{figure}

@@

    	\item People engage in a more social and human-like way with a robotic pet compared to a functional robot and a tablet computer, therefore there will be a higher amount of anthropomorphic language in the AIBO and Roomba forum, compared to the iPad forum.

    	\item As anthropomorphism can be seen as a certain kind of emotional / social relation to an artifact, there will be more anthropomorphic language in forum posts in which the author describes a relation to the device, compared to discussing technological aspects or how the device is used.
\end{enumerate}

@@

\begin{itemize}
    \item \textit{Life-likeness}, such as being alive or having (parts of) a
        body, \eg \textit{``she can be reborn since you copied her memory
        stick''};

    \item \textit{Emotional states, empathy} or the technology having a
        \textit{feeling}, \eg \textit{``oh no, poor AIBO''};

    \item \textit{Gender, personality} or the technology having an
        \textit{intention}, \eg Roomba \textit{``seems to like to hang around
        under the sofa''};

    \item When the author gave the technology a \textit{name}, \eg
        \textit{``whenever I show Java his pink ball''};

    \item Being \textit{socially integrated}, such as being a family member, \eg
        \textit{``I'm considering adding a Roomba to the family''};

    \item \textit{Metaphorical ways}, \eg Roomba \textit{``sings its victory
        song when it finishes and docks''}.

\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		WHY DO PEOPLE ANTHROPOMORPHIZE?
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph*{Developmental psychology research on the attribution of animacy and intention\\ \\}

@@

\paragraph*{Cognitive science research and the mental model theory (Theory of Mind)\\ \\}

@@

\paragraph*{Social psychology research: a 3-factor theory of anthropomorphism\\ \\}
\label{sec:psychological-factors}

@@

@@

\paragraph{Why is anthropomorphism likely to change over time?}
\hl{maybe we need to move this paragraph to PART 3}

\subsection{A Model of the Dynamics of Anthropomorphism}

>>>>

Figure~\ref{fig:roomba-perception} gives a first perspective on how naive
users perceive robots over a long period of time. Here, \emph{Roomba} vacuum cleaning robots
were given to nine households for a 6-months period. The aim was to explore how
people use and perceive the robot over time, and how they generally live with
the Roomba in their home. At various time points, users were asked to rate
their perception of different features of the robot on a 7-point rating scale. As
Figure~\ref{fig:roomba-perception} shows, some features remained stable over time
(\emph{ease of use} and \emph{fun}), and others decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one does not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on his/her life. Even if one can argue that the Roomba is not designed
to foster interaction, the question is raised: how to sustain engagement between
a robot and a human over a long period of time? And, as a pre-requisite, do we
actually understand the psychological phenomenon, and can we effectively account
for the dynamics of these interactions?

@@

@@

@@

%    \begin{tikzpicture}[scale=0.5, y = 2cm]
%    %axis
%    \draw (0,0) -- coordinate (x axis mid) (20,0);
%    \draw (0,0) -- coordinate (y axis mid) (0,7);
%    %ticks
%    \foreach \x in {0,4,...,20}
%        \draw (\x,1pt) -- (\x,-3pt);
%    \foreach \y in {1,...,7} 
%        \draw[black!20] (20,\y) -- (-3pt,\y);
%    \foreach \y in {1,...,7} 
%        \draw (-3pt, \y) node[anchor=east] {\y}; 
%    %labels
%    \node[below=0.8cm] at (x axis mid) {People's perception (n = 15)};
%    \node at (2, -0.4) {\tiny expectations};
%    \node[align=center] at (6, -0.4) {\tiny after\\ \tiny first usage};
%    \node[align=center] at (10, -0.4) {\tiny after\\ \tiny 2 weeks};
%    \node[align=center] at (14, -0.4) {\tiny after\\ \tiny 2 months};
%    \node[align=center] at (18, -0.4) {\tiny after\\ \tiny 6 months};
%
%    \draw[dashed, black!50] (6,0) -- (6,7) node[sloped, below, pos=0.25] {\tiny novelty effect};
%
%    \draw[ultra thick, red!50] (2,5.6) -- (6,6.1) -- (10,6.3) -- (14,6.3) -- (18,6.1) node at (15,6.5) {\tiny \it ease of use};
%    \draw[ultra thick, purple] (2,4.9) -- (6,5.2) -- (10,5.1) -- (14,4.5) -- (18,4.2) node at (3,4.7) {\tiny \it usefulness};
%    \draw[ultra thick, blue!50] (2,4) -- (6,4.1) -- (10,4.6) -- (14,4.1) -- (18,4.3) node at (16,4) {\tiny \it fun};
%    \draw[ultra thick, green!50] (2,3.9) -- (6,4.1) -- (10,4.2) -- (14,3.8) -- (18,3.3) node at (19,3.5) {\tiny \it intelligence};
%    \draw[ultra thick, orange] (2,3.5) -- (6,4.2) -- (10,4) -- (14,3.2) -- (18,2.8) node at (14,2.7) {\tiny \it impact};
%
%    \draw[ultra thick, dashed] (2,5.2) -- (6,5.2) -- (10,5) -- (14,4.9) -- (18,4.5) node at (14,5.3) {\tiny \it overall};
%
%    \end{tikzpicture}

@@

In accordance with previous work we also found relatively low rates of users
treating and perceiving their vacuum cleaning robot as a living (or human-like)
entity \citep{sung_housewives_2008}. We interpret that it might be the pure
functional purpose and non human-like form of the Roomba that prevented people
from anthropomorphizing it. Still, we observed some initial anthropomorphic
effects, both reflected in interaction with and perception of the robot.
However, most of the anthropomorphic effects did not last long. One fairly
stable human-like interaction with the robot was \textit{talking to the robot},
despite the fact that the robot was not able to recognize and respond to verbal
statements.

@@

%\begin{figure}
%    \centering
%    \includegraphics[width=0.7\columnwidth]{aibo-roomba-ipad_2.jpg}
%    \caption{The Sony AIBO robotic dog (left), the iRobot Roomba vacuum-cleaning robot (middle), and the Apple Inc. iPad tablet computer (right) were compared in terms of how far users of these devices use anthropomorphic expressions in discussions about their technology / robot in online discussion forums.}
%    \label{fig:aibo-roomba-ipad}
%\end{figure}

@@

\begin{enumerate}
    	\item People engage in a more social and human-like way with a robotic pet compared to a functional robot and a tablet computer, therefore there will be a higher amount of anthropomorphic language in the AIBO and Roomba forum, compared to the iPad forum.
    	\item As anthropomorphism can be seen as a certain kind of emotional / social relation to an artifact, there will be more anthropomorphic language in forum posts in which the author describes a relation to the device, compared to discussing technological aspects or how the device is used.
\end{enumerate}

@@

\begin{itemize}
    \item \textit{Life-likeness}, such as being alive or having (parts of) a
        body, \eg \textit{``she can be reborn since you copied her memory
        stick''};
    \item \textit{Emotional states, empathy} or the technology having a
        \textit{feeling}, \eg \textit{``oh no, poor AIBO''};
    \item \textit{Gender, personality} or the technology having an
        \textit{intention}, \eg Roomba \textit{``seems to like to hang around
        under the sofa''};
    \item When the author gave the technology a \textit{name}, \eg
        \textit{``whenever I show Java his pink ball''};
    \item Being \textit{socially integrated}, such as being a family member, \eg
        \textit{``I'm considering adding a Roomba to the family''};
    \item \textit{Metaphorical ways}, \eg Roomba \textit{``sings its victory
        song when it finishes and docks''}.
\end{itemize}

@@

@@

\paragraph*{Developmental psychology research on the attribution of animacy and intention\\}

@@

\paragraph*{Cognitive science research and the mental model theory (Theory of Mind)\\}

@@

\paragraph*{Social psychology research: a 3-factor theory of anthropomorphism\\}
\label{sec:psychological-factors}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		ANTHROPOMORPHISM AS A MULTI-FACTOR PHENOMENON
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@
